### 2024-09-22 cognitive-bias.qmd
20240922 - omission bias and regret

*Omission bias* is the tendency to prefer inaction over action.
For instance, consider a manager who is on the fence to drop an underperforming player and to pick up someone else to start.
If the manager does not drop the player and the player scores few points, the manager is likely to be less critical of their decision than if they had picked up another player who then scored the same (low) number of points.

@@

A *sunk cost* is a cost (e.g., in money, time, or effort) that has already been incurred and cannot be recovered.
For instance, if a person orders an expensive meal at a restaurant, the order is a sunk cost.
The *sunk cost fallacy* is the tendency to continue an endeavor when there is a sunk cost.
For instance, when ordering the expensive meal at the restaurant, a person may over-eat so that they feel that they eat their money's worth of food.

>>>>

*Omission bias* is the tendency to prefer inaction over action, or to prefer the default option over changes from the default option [@Kahneman2011].
For instance, consider a manager who is on the fence to drop an underperforming player and to pick up someone else to start.
If the manager does not drop the player and the player scores few points, the manager is likely to be less critical and regretful of their decision than if they had picked up another player who then scored the same (low) number of points.
Thus, a person's decisions may be influenced by the anticipation of regret so as to keep the status quo, avoid losses, and minimize the expectation of regret—such as choosing the same order each time at a restaurant to avoid regret [@Kahneman2011].
However, people may anticipate more regret than they actually experience, because people tend to underestimate the protective impact of the psychological defenses they will deploy against regret [@Kahneman2011].
In general, @Kahneman2011 recommends not to put too much weight on potential future regret when making decisions: "even if you have some [regret], it will hurt less than you now think." (p. 352).

@@

A *sunk cost* is a cost (e.g., in money, time, or effort) that has already been incurred and cannot be recovered.
For instance, if a person orders an expensive meal at a restaurant, the order is a sunk cost.
The *sunk cost fallacy* is the tendency to continue an endeavor when there is a sunk cost.
For instance, when ordering the expensive meal at the restaurant, a person may over-eat so that they feel that they eat their money's worth of food.
The sunk cost fallacy is an example of the commitment bias, an escalation of commitment by which a person continues to commit additional time, effort, or resources to a failing project.
It is important to be willing to identify and terminate failing projects to cut one's losses.
